The impact of gas bulk rotation on the Lyman $\alpha$ line

### 2014-06-03 paper/referee-reply.tex

\textbf{Thanks very much for this recommendation we made this new plot and we do find variations in the morphology of the line due to variations in the viewing angle. We have restructure the paper this new plots now correspond to figure 2 \& 3 in the paper.  Also we emphasize this results in the results, discussion and conclusion sections.\\}

What about the variation of lya escape fraction with viewing angle ? I propose that in the beginning of Sect.3 Results, you first emphasize that, maybe counter-intuitively, you did not find any variation of the Lya properties with viewing angle, so you will present first angle average lya properties, and you will come back to the anisotropy problematic only at the end on the paper.\\

\textbf{As escape fraction depends on the number of scatterings and this is invariant for different viewing angles the escape fraction does not change with the viewing angle. (MAKE A PLOT AND EXPLAIN THIS BETTER)
}

\begin{figure}
\includegraphics[scale=0.5]{escapefractionvstheta.png}
\caption{Escape Fraction as a function of $\mu$}
\end{figure}

@@

\subsection*{End of paragraph 3.3}

One sentence is not clear : we see that the escape fraction does not increase significantly from $\tau$ = 105 to $\tau = 10^{6}$. This counter-intuitive result.... It sounds like you were expecting a strong increase... A decrease is expected from $\tau$ = 105 to $\tau$ = 106, not an increase, but indeed on the graphe we can see an unexpected (small) increase. I do not understand the explanation for this behaviour.
(Mark)

\subsection*{Fig8}

Referred to as Figure 7 in the text (paragraph 3.4). To my mind, this figure illustrates the main result of your study, it has to be more demonstrative. On this figure, the numerical noise seems very big compare to the small number of bins, bigger than on the spectra from Fig1, how is it possible ? Could you re-do this figure with more photons, maybe with more bins ? Could you plot an histogram instead of broken lines, to ease the reading ? In the text, you write anisotropy induced by rotation is at the 3$\%$ level, how do to estimate this number? I think that the noise is at $3\%$ level, and the distribution is compatible with ’flat’. In the corresponding text, you first say that $F(\mu)$ does not depend on Vmax. Again, if real, this is an important result, that you must advertise more. Please show it on this Fig, by comparing F($\mu$) without rotation (in red, for example) to the other distributions that you get with different values of Vmax. Then, you mention that for high optical depth values, $F(\mu)$ (not the variations of $\mu$) can vary up to $15\%$. First, what do you call high optical depth ? your highest setup is $\tau = 10^{7}$, not extremely high for lyman alpha ”standards’, corresponding to NHI ∼ 1020 cm−2. So, if you see a trend with optical depth, I encourage you to check higher optical depth regimes, corresponding better to galaxy scales. If you can probe that anisotropy induced by rotation is rising with optical depth, this would be an interesting result.
If higher optical depth regimes lead to anisotropic escape, could you check the effect of anisotropy on the lya spectral shape ?

### 2014-06-04 paper/referee-reply.tex

\documentclass[12pt]{article}
\usepackage[margin=1.0in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{graphicx}

\title{Response to the first report on the article draft \emph{The Impact of Gas Bulk Rotation on the Lyman-$\alpha$ line.}}
\author{Juan N. Garavito-Camargo, Jaime E. Forero-Romero, Mark Disjkstra}
\date{\today}

@@

\begin{document}

\maketitle
We thank the referee for a detailed report. In this occation the
report by the referee motivated us to explore a very important aspect
of our results that we had overlooked.

At the moment of addressing the comments in the report, we
prepared a plot suggested by the referee. The plot showed a clear
dependence of the line morphology with the viewing angle, a result
that we had deemed negligeable after measuring that the integrated
flux did not have any dependence on the viewing angle

The correct interpretation is that, as the referee expected, there is
a strong effect on the morphology.

This completely changed the focus of the previous version of the
paper. We have now completely rewritten the Results and Discussion
section in order to highlight that result. We have also completeley dropped
sections (e.g. the asymetrical emission part) that obfuscate the main
result of the paper.

In this document we address one by one the questions in the
report. However, we kindly ask the referee to revise directly the new
version of the paper, given the substantial changes from the first
version.

In what follows the comments by the referee are boldfaced.

Best regards, \\

The Authors\\

@@

\section{Global Comments}

\subsection{Variation with the viewing angle}

{\bf The problematic of anisotropy is only studied from Sect3.4, the former
sections consider global quantities (spectral shapes, escape
fractions, etc...). This is ok only if there is NO anisotropy induced
by rotation, which is not obvious, a priori. You should make it clear
from the beginning, telling that you will investigate anisotropy at
the end of the paper only, because you checked that lya properties are
isotropic, at least in the range of parameters that you investigated
so far.}

We have decide to study in this paper the case in which there is No anisotropy. The part concerning the off-center emission is going to be presented in a later study.

{\bf As explained in sect 3.4, rotation kills the spherical symmetry of your problem, and the rotation axis defines a preferential direction. We could expect some variation of the emerging flux, the lyman-alpha escape fraction, and the spectral shape, with viewing angle. From Fig8, it seems that the flux is the same in all directions for spherical distributions of sources. This important result is not enough emphasized.}

We do not find the flux variation with the viewing angle and now as
highlighted in the new text we emphasize more this new result, also we
made a new plot Figure 5 in the paper, showing this results for both
distributions (homogeneous/central) and for optical depth
$\tau_{H}=10^{5}$ we choose these models because this are the ones  that are more sensitive to rotation.

{\bf Do you see any difference in the spectral shape with viewing angle? To illustrate this point, it would be interesting to build a 2D plot as you did on Fig 5, but replace Nb scat in ordinates by the viewing angle $\mu$. We could immediatly see if there is an evolution of the shape with viewing angle or not. If you find NO evolution of the Lya shape with viewing angle, I think that it is an interesting counter-intuitive result, that you should advertise more.}

Thanks very much for this recommendation we made this new plot and we do find variations in the morphology of the line due to variations in the viewing angle. We have restructure the paper this new plots now correspond to figure 2 \& 3 in the paper.  Also we emphasize this results in the results, discussion and conclusion sections.

{\bf What about the variation of lya escape fraction with viewing angle ? I propose that in the beginning of Sect.3 Results, you first emphasize that, maybe counter-intuitively, you did not find any variation of the Lya properties with viewing angle, so you will present first angle average lya properties, and you will come back to the anisotropy problematic only at the end on the paper.}

As escape fraction depends on the number of scatterings and this is invariant for different viewing angles the escape fraction does not change with the viewing angle. (MAKE A PLOT AND EXPLAIN THIS BETTER)

@@

We want to study the pure effect of rotation on the spectra, off course it would
be interesting to study the effect of rotation on systems with
outflows, this paper is under preparation by M.C Remolina-Gutierrez et
al.

\section*{Details}

\subsection*{Introduction}

{\bf With Orsi et al 2012, please cite also Garel et al 2012. With Zheng \& Wallace 2013, please cite also Behrens et al 2014.}

Added.

\subsection*{Fig.1}

{\bf I guess that the spectra presented in Fig1 are integrated over all directions, right ? You should describe explicitly how you build them. You could skip the x notation in absciss, as it is not used in the discussion, whereas the velocity is used to compare to FWHM, on Fig2.}

Now this figure is the Fig.4 in the paper. what we did was to fix a viewing angle in this case $\theta= \pi/2$ and make the histogram of the photons frequency

{\bf Description about how we build Fig.1}

 (Done)
{\bf Fig 1. remove x notation}

(Done)

@@

\subsection*{Fig.2}

{\bf Fig.2 Can you explain how you measure the FWHM of a double-peaked profile ? Do you fit it with a gaussian ?}

We e interpolate points to the line profile and then  we measure the
width by  the closest points to the half maximum in our histogram.

@@

\subsection*{Fig.3}

{\bf Do you have an idea why the (central source, intermediate optical depth) case with Vmax=300 has a single plateau instead of 2 peaks ? Do you find this with the two codes ?}

After close inspection in this figure the line was composed by two
peaks, that looked like a plateau with the resolution used to
construct the figure.

@@

\subsection*{Fig4 and 6}

{\bf This is a surprising result that the number of scatterings (escape fraction) stays constant as the rotation velocity increases, for a central source, whereas the global spectral shape is altered. Did you try with higher/extreme values of Vmax (=1000 km/s, even if not physically motivated) ? Do you believe that the number of scatterings decreases with very high values, or that it is independant of the rotation velocity ? Is the escape fraction from a dusty rotating cloud with central source independant of the rotation velocity ?}

\subsection*{Fig5}

{\bf This is a very nice figure ! Looking at the top right panel, with
  its “photosphere”, I’m surprised that the distribution of Nscatt is
  bipolar, I would have expected a smooth transition between the 2
  regimes. Do you have a physical explanation why photons escape after
  either (less than 10) or (more than 1000) scatterings, and none
  escape with 100 scatt ? To test the ’photosphere’ assumption, you
  could also do the radiation transfer in a cloud with the
  distribution of sources only in an inner sphere, and another
  experienment with sources only in an outer shell. If I understood
  corectly, you would expect thant the bottom spot on the top right
  panel is made of photons from the outer shell, and the uppper spot
  from photons emitted inside the cloud.}

@@

\begin{figure}
  \includegraphics[scale=0.4]{Histogram2dNscattVSRadius.png}
\end{figure}

\subsection*{End of paragraph 3.3}

{\bf One sentence is not clear : we see that the escape fraction does not increase significantly from $\tau$ = 105 to $\tau = 10^{6}$. This counter-intuitive result.... It sounds like you were expecting a strong increase... A decrease is expected from $\tau$ = 105 to $\tau$ = 106, not an increase, but indeed on the graphe we can see an unexpected (small) increase. I do not understand the explanation for this behaviour.}

(Mark)

@@

\subsection*{Fig8}

{\bf Referred to as Figure 7 in the text (paragraph 3.4). To my mind,
  this figure illustrates the main result of your study, it has to be
  more demonstrative. On this figure, the numerical noise seems very
  big compare to the small number of bins, bigger than on the spectra
  from Fig1, how is it possible ? Could you re-do this figure with
  more photons, maybe with more bins ? Could you plot an histogram
  instead of broken lines, to ease the reading ? In the text, you
  write anisotropy induced by rotation is at the 3$\%$ level, how do
  to estimate this number? I think that the noise is at $3\%$ level,
  and the distribution is compatible with ’flat’. In the corresponding
  text, you first say that $F(\mu)$ does not depend on Vmax. Again, if
  real, this is an important result, that you must advertise
  more. Please show it on this Fig, by comparing F($\mu$) without rotation
  (in red, for example) to the other distributions that you get with
  different values of Vmax. Then, you mention that for high optical
  depth values, $F(\mu)$ (not the variations of $\mu$) can vary up to
  $15\%$. First, what do you call high optical depth ? your highest
  setup is $\tau$= $10^7$, not extremely high for lyman alpha ”standards’,
  corresponding to NHI $\sim$10$^{20}$ cm$^{-2}$. So, if you see a trend with
  optical depth, I encourage you to check higher optical depth
  regimes, corresponding better to galaxy scales. If you can probe
  that anisotropy induced by rotation is rising with optical depth,
  this would be an interesting result. If higher optical depth regimes
  lead to anisotropic escape, could you check the effet of anisotropy
  on the lya spectral shape?}

### 2014-06-04 paper/ms.tex
All figures in PDF and PNG
minor updates

We present results of radiative transfer calculations to measure the
impact of gas bulk rotation on the morphology of the Lyman $\alpha$ 
emission line in distant galaxies. We model a galaxy as a sphere with
an homogeneous mixture of dust and hydrogen at a constant
temperature. These spheres have a solid-body rotation with maximum
velocities in the range $0-300$ \kms and neutral hydrogen optical
depths in the range $\tau_{\rm H}=10^{5}-10^{7}$. We consider cases 
of a single central \lya source, and a uniform distribution of sources. 
Our main result is that rotation introduces a dependence of the
line morphology with  viewing angle. Observations with a line of sight
parallel to the rotation axis yield line morphologies similar to the
static case. The greatest difference is observed for lines of sight
perpendicular to the rotation axis. In this case the intensity at the line
center increases and the line width increase with rotational velocity.  For
homogeneously distributed sources the line becomes single peaked  at
rotational velocities larger than half the line width in the static
case. However, rotation does not not induce any significant anisotropy
in the total line intensity nor in the escape fraction. 
\end{abstract}

@@

Figure \ref{fig:muhisto} shows the results for a selection of models
with $\tau_{\rm H}=10^{5}$, different rotational velocities and both
for the central and homogeneous source distributions. The
distributions do not show any trend, beyond some fluctuations below
the $2\%$ level. This result implies that the integrated flux is
constant for all viewing angles and all the possible models.

@@

\begin{figure*}
\begin{center}
  \includegraphics[width=0.40\textwidth]{f6.eps}
  \includegraphics[width=0.40\textwidth]{f7.eps}
\end{center}
  \caption{Effects of rotation on the escape fraction. Left
    panel, escape fraction as a function of rotational velocity. All
    these models have $\tau_{a}=1$. The continuous (dashed) lines
    correspond to central (homogeneous) models. Right panel, escape
    fraction as a    function of the product $(a\tau_{\rm
      H})^{1/3}\tau_{A}$. The    analytic solution for the infinite
    slab is shown as a continuous line. Different lines correspond
    to different rotational velocities for the homogeneous models.     
    \label{fig:escape_fraction}} 
\end{figure*}

@@

\begin{figure*}
\begin{center}
    \includegraphics[width=0.45\textwidth]{nscattvsmu.eps}
    \includegraphics[width=0.45\textwidth]{f4.eps}
\end{center}
\caption{Logarithm of the average number of scatterings as function of
  $\mu$ (left panel) and the maximum rotational velocity $V_{\rm
    max}$. The left panel shows the behaviour for $\tau=10^{5}$ and
  $V_{max}=300$\kms as a function of $|\cos\theta|$; the independence of
  $N_{\rm scatt}$ with $\mu$ is present in all models. 
  In the right panel the continuous (dashed) lines represent an
  central (homogeneous) distribution of sources; there is a clear
  decrease in the number of scatterings for the homogenously
  distributed sources. 
\label{fig:Nscatt} }    
\end{figure*}

\begin{figure*}
\begin{center}
  \includegraphics[width=0.48\textwidth]{f5_1.eps}
  \includegraphics[width=0.48\textwidth]{f5_2.eps}
  \includegraphics[width=0.48\textwidth]{f5_3.eps}
  \includegraphics[width=0.48\textwidth]{f5_4.eps}    
\end{center}
    \caption{2D histogram of $N_{\rm scatt}$ vs $x$. The upper (lower) panels
      show the homogeneous (central) source distribution. Left panels 
      corresponds to the static case and right panels correspond to
      $V_{\rm max}=300\kms$. The color scale is logarithmic on the
      number of photons with given values of $N_{\rm scatt}$ and
      $x$. \label{fig:Nscatt2D}. A results that stands out is in the upper
      right panel, showing that in homogeneously distributed sources a
      large number of photons can escape escape after a few
      scatterings, $N_{\rm scatt}<10$.}   
\end{figure*}

@@

In contrast, we found two quantities that are invariant with respect
to the viewing angle and the rotational velocity: the integrated flux and the escape
fraction.  The exception are the models with homogeneously distributed
sources, which present an escape fraction dependent on the rotational
velocity, but still independent of the viewing angle. We showed that
these results could be interpreted in terms of the average number of
scatterings for the photons.

>>>>

We present results of radiative transfer calculations to measure the
impact of gas bulk rotation on the morphology of the Lyman $\alpha$ 
emission line in distant galaxies. We model a galaxy as a sphere with
an homogeneous mixture of dust and hydrogen at a constant
temperature. These spheres have a solid-body rotation with maximum
velocities in the range $0-300$ \kms and neutral hydrogen optical
depths in the range $\tau_{\rm H}=10^{5}-10^{7}$. We consider cases 
of a single central \lya source, and a uniform distribution of sources. 
Our main result is that rotation introduces a dependence of the
line morphology with  viewing angle. Observations with a line of sight
parallel to the rotation axis yield line morphologies similar to the
static case. The greatest difference is observed for lines of sight
perpendicular to the rotation axis. In this case the intensity at the line
center increases and the line width increase with rotational velocity.  For
homogeneously distributed sources the line becomes single peaked  at
rotational velocities larger than half the line width in the static
case. Notably, we find that rotation does not not induce any
significant spatial anisotropy in the integrated line flux nor in the escape
fraction.  However, we do find in the homogeneous models a clear
increase of the escape fraction with increasing rotational velocity.
\end{abstract}

@@

Figure \ref{fig:muhisto} shows the results for a selection of models
with $\tau_{\rm H}=10^{5}$, different rotational velocities and both
for the central and homogeneous source distributions. The
distributions do not show any trend, beyond some fluctuations below
the $2\%$ level.

This is a remarkable result. In spite of that the rotation axis
defines preferential direction, the integrated flux is the same for
all viewing angles and all the possible models in our spherical
distributions.

@@

\begin{figure*}
\begin{center}
  \includegraphics[width=0.40\textwidth]{f10.pdf}
  \includegraphics[width=0.40\textwidth]{f11.pdf}
\end{center}
  \caption{Effects of rotation on the escape fraction. Left
    panel, escape fraction as a function of rotational velocity. All
    these models have $\tau_{a}=1$. The continuous (dashed) lines
    correspond to central (homogeneous) models. Right panel, escape
    fraction as a    function of the product $(a\tau_{\rm
      H})^{1/3}\tau_{A}$. The    analytic solution for the infinite
    slab is shown as a continuous line. Different lines correspond
    to different rotational velocities for the homogeneous models.     
    \label{fig:escape_fraction}} 
\end{figure*}

@@

\begin{figure*}
\begin{center}
    \includegraphics[width=0.45\textwidth]{f12.pdf}
    \includegraphics[width=0.45\textwidth]{f13.pdf}
\end{center}
\caption{Logarithm of the average number of scatterings as function of
  $\mu$ (left panel) and the maximum rotational velocity $V_{\rm
    max}$. The left panel shows the behaviour for $\tau=10^{5}$ and
  $V_{max}=300$\kms as a function of $|\cos\theta|$; the independence of
  $N_{\rm scatt}$ with $\mu$ is present in all models. 
  In the right panel the continuous (dashed) lines represent an
  central (homogeneous) distribution of sources; there is a clear
  decrease in the number of scatterings for the homogenously
  distributed sources. 
\label{fig:Nscatt} }    
\end{figure*}

\begin{figure*}
\begin{center}
  \includegraphics[width=0.48\textwidth]{f14_1.pdf}
  \includegraphics[width=0.48\textwidth]{f14_2.pdf}
  \includegraphics[width=0.48\textwidth]{f14_3.pdf}
  \includegraphics[width=0.48\textwidth]{f14_4.pdf}    
\end{center}
    \caption{2D histogram of $N_{\rm scatt}$ vs $x$. The upper (lower) panels
      show the homogeneous (central) source distribution. Left panels 
      corresponds to the static case and right panels correspond to
      $V_{\rm max}=300\kms$. The color scale is logarithmic on the
      number of photons with given values of $N_{\rm scatt}$ and
      $x$. \label{fig:Nscatt2D}. A results that stands out is in the upper
      right panel, showing that in homogeneously distributed sources a
      large number of photons can escape escape after a few
      scatterings, $N_{\rm scatt}<10$.}   
\end{figure*}

@@

Remarkably, we find two quantities that are invariant with respect
to the viewing angle and the rotational velocity: the integrated flux and the escape
fraction.  The exception is the models with homogeneously distributed
sources, which present an escape fraction dependent on the rotational
velocity, but still independent of the viewing angle. We showed that
these results could be interpreted in terms of the average number of
scatterings for the photons.
